---
title: "Cytoreason Single-cell Analysis Pipeline"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = FALSE)
library(plyr)
library(cytoreason)

devtools::load_all("~/capsule/code/cytoreason.single.cell/")
devtools::load_all("~/capsule/code/cytoreason.cc.client/")
```

# Overview

* Requirements:
  - a workflow ID from an export single-cell dataset, as output by `sc_api_save_data()`
  - a configuration list that describes the dataset and the analysis models to fit
  - a supported disease model short key, i.e. available in `get_disease_model_metadata()`
  
* The pipeline proceeds as follows:
  - 

# Configuration
```{r config}
config <- sc_api_configuration(
  # asset_id [mandatory]: workflow ID of an exported single-cell dataset as 
  # output by `sc_api_save_data()`
  asset_id = "wf-0ccf093f95",
  #asset_id = 'wf-a8e301eaec',
  cell_annotation = "biologist_annotation", # or: "final_2023"
  # experiment_id [optional]: experiment identifier.
  # If not provided, the default is to use the asset_id.
  # It does not need to be a GSE ID. 
  # Sample annotations are looked up in the Currator using this key, but the 
  # pipeline will not fail if it does not find any
  experiment_id = "GSE231993",
  
  # platform_id [optional]: platform ID, if not provided it will be "sc-rnaseq" 
  # Note however that the platform ID in the text files used by the platform API will  
  # always be set to "sc-rnaseq".
  platform_id = "GPL18573",
  
  # biosample_id [optional]: a string that indicates the name of the cell annotation
  # variable (in the cell metadata), which holds the bio-sample identifier, 
  # i.e. the key that is used to aggregate the single-cell data into a 
  # pseudobulk data
  biosample_id = "sample",
  # biosample_id = "donor_id",
  
  # sample_annotations [optional]:
  sample_annotations = list(
    # condition: values must be from the DOID ontology used in the Curator.
    # Here we can remap invalid values in the original data into valid ontology terms.
    # Values that are not remapped remain untouched.
    # Format: list(<original_variable_name> = c(new_value1 = old_value1, new_value2 = old_value2))
    condition = list(
      condition = c(control = "healthy",
                  "ulcerative colitis" = "ulcerative colitis")
    ),
    # tissue: values must be from the BTO ontology used in the Curator.
    # Here we can remap invalid values in the original data into valid ontology terms 
    # See example on 'condition' for format description
    tissue = list(
      tissue = c(
        # "colon" = "Normal ascending colon",
         "colon" = "colon ascendens",
         "colon" = "colon" 
      )
    ),
    # 
    sample_classification = list(
      sample_classification = c("Non inflamed" = "Non inflamed", 
               "Inflamed" = "Inflamed", 
               Normal = "Normal")
    ),
    # directly use column `donor_id` as `subject_id`
    subject_id = "donor_id"
    # # a custom variable as a cross-product of 2 other variables
    # custom_variable = list(
    #   type = c("NonI", "Infl", "Heal"),
    #   disease = c(healthy = "normal",
    #               "CD" = "Crohn disease")
    # )
  ),
  term_metadata = list(

     UC = list(
        DZ_vs_HC = list(
          contrast_type = "disease_vs_control"
          )
        ),
    
     I_vs_NI = list(
       # will only export this contrast term within this comparison
       I_vs_NI = list(contrast_type = "inflamed_vs_non_inflamed")
     ),
      
     UC_inf = list(
       # will only export this contrast term within this comparison
       DZ_inf_vs_HC = list(contrast_type = "inflamed_vs_healthy")
     )
  ),
  analysis_model = list(

    UC = list(
      group = list(
        # this will fit the same term as use-case 2, but label it as "DZ_vs_HC"
        condition = c(HC = "control", DZ = "ulcerative colitis")
      )#,
        # pairing = NULL,
        #covariates = c("assay","layer")
      #covariates = "layer"
    ),
      I_vs_NI = list(
        # group variable: use `sample_classification`, while filtering
        # and enforcing level order
        group = list(
          sample_classification = c(NI = "Non inflamed", I = "Inflamed")
        )#,
        #,
        # pairing= "donor_id",
         #covariates = c("assay","layer")
        #covariates = "layer"
      ),
        UC_inf = list(
        # group variable: use `sample_classification`, while filtering
        # and enforcing level order
        group = list(
          sample_classification = c(HC = "Normal", DZ_inf = "Inflamed")
        )#,
        # pairing= "donor_id",
         #covariates = c("assay","layer")
        #covariates = "layer"
      )
  ),
  parameters = list(
    model=list(
      services = c("gx_diff", "ct_test"),
      gx_diff=list(min_sample_size=15),
      ct_test=list()
    )
  )
)


```

```{r}
# check the configuration locally
sc_validate_dataset_configuration(config)
# check that a valid dataset can be loaded from the config
pb_object <- prepare_single_cell_dataset(config)
# check a particular analysis model (if failing)
obj <- prepare_single_cell_dataset(config, validate = FALSE)
str(build_analysis_model_metadata(config$analysis_model$UC, obj))
```

## Available model keys
```{r model_keys, echo = FALSE}
knitr::kable(ldply(get_disease_model_metadata(), .id = "model_key", as.data.frame))
```

# Running the pipeline
```{r}
# launch piepline
sc_api_run_analysis("UC-CO", config, tags = list(message = "fixed groups names"))
# [Mon Jan 15 13:00:10 2024] run_analysis - wf-2b0097bbd3
# wf-d48a85fa3f after removing features[features==0] <- 0.01 from DA code
# [Sun Jan 21 15:18:52 2024] run_analysis - wf-424bee66dd - fixed unclassified B cells typo
# [Wed Jan 24 12:11:00 2024] run_analysis - wf-7445c55bf9
```

```{r}
model_md <- get_disease_model_metadata("UC-CO")
```

```{r}
model_key = "UC-CO"
test_run <- .worker__sc_analysis_pipeline(model_key, config)
```

# BQ upload ----
```{r}

combine_ds <- sc_api_save_analysis(c('wf-37842c37af','wf-7445c55bf9'), dataset_name = "p00_ulcerative_colitis")

command <- paste("python /app/cytobigquery/exec_service.py single_cell", wf_id = "wf-7445c55bf9", dataset_name = "p00_ulcerative_colitis", "--verbose -sc -ti=eu.gcr.io/cytoreason/cd-py-bigquery:develop_latest -tm=124Mi")
res <-
  run_command_dist(
    command,
    image = "eu.gcr.io/cytoreason/cd-py-bigquery:develop_latest",
    memory_request = "25Gi",
    force_execution = FALSE,
    replace_image_tags = TRUE
    )
#wf-b785ca238f
# wf-5a5932eacc - upload of wf-d48a85fa3f
#wf-665e6cbf73 - upload of wf-424bee66dd
#wf-d3f28e4147 - upload of wf-7445c55bf9
```



# Troubleshooting

## Tissue or disease annoations
```{r ontology}
sc_obj <- get_single_cell_dataset(asset_id = "wf-0511d85469", cell_annotation= "biologist_annotation")
# look at which annotation is invalid
table(sc_obj$tissue)
setdiff(sc_obj$tissue, get_ontology_tissue_terms())
# match them in the ontology
grep("small intestine", get_ontology_tissue_terms(), value = TRUE)
grep("lamina propria", get_ontology_tissue_terms(), value = TRUE)
grep("ileal|ileum", get_ontology_tissue_terms(), value = TRUE)
grep("Crohn", get_ontology_disease_terms(), value = TRUE)
```